---
title: "Add Variables Using External Data - Programs"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

------------------------------------------------------------------------

We add columns to the dataset for **each USDA broadband program**. Each program column will contain a program identifier code for each property that is within the eligibility area for that program. We create these columns by finding the geographic intersection of the program's eligibility areas (as specified in their respective shapefiles) with the coordinates of the corelogic properties. The final output is a new table with a column for each broadband program. If a property is physically within the eligibility area for one of the programs, then a program identifier is entered in that program's column.

## Load libraries and functions

```{r}
source("../R/functions.R")
library(data.table)
library(RPostgreSQL)
library(sf)
library(glue)
library(magrittr)
```

### Set Geo Data Directory

```{r, cache=TRUE}
# {r, cache=TRUE}
geodatadir <- "../../../data/projects_data/usda/bb/original/geo"
```

### Get list of all U.S. counties

For use with program data that don't have county fips codes. Convert list to sf.

```{r, cache=TRUE}
# {r, cache=TRUE}
counties_us <- tigris::counties() %>%
  sf::st_transform(4326)
```

## CC Program Eligible Properties

#### Load broadband program area shapefiles

```{r, cache=TRUE}
# {r, cache=TRUE}
map_program_area_cc <- st_read(file.path(geodatadir, "CC 2013_2019_83 04272020.shp")) %>%
  sf::st_transform(4326)
```

#### Create list of county FIPS codes that are within the program area for the database query

```{r}
# {r}
counties_cc <- paste0("'", paste(unique(strsplit(paste(map_program_area_cc$FIPS, collapse = ", "), ", ")[[1]]), collapse = "','"), "'")

counties_cc_sql <- glue_sql(counties_cc)
```

#### Get property records for the program related counties from the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)
```

```{sql}
-- {sql connection=con, output.var="qry_res"}
SELECT geoid_cnty, p_id_iris_frmtd, property_centroid_longitude, property_centroid_latitude FROM corelogic_usda.current_tax_200627_latest_all_add_vars 
WHERE property_centroid_longitude IS NOT NULL 
AND geoid_cnty IN (?counties_cc_sql);
```

```{r}
# {r}
DBI::dbDisconnect(con)
```

#### Get intersection between program area shapefile and the selected properties

First convert database query result to an sf object.

```{r}
# {r}
db_rows_cc <- st_as_sf(qry_res, coords = c("property_centroid_longitude", "property_centroid_latitude"), crs = 4326)

int <- sf::st_intersects(db_rows_cc, map_program_area_cc)
int2 <- as.integer(as.character(int))
```

#### Add RUSID to intersecting properties

```{r}
# {r}
db_rows_cc$program <- "CC"
db_rows_cc$rusid <- as.character(map_program_area_cc$RUSID[unlist(int2)])
db_rows_cc <- db_rows_cc[!is.na(db_rows_cc$rusid),]
```

## RC Program Eligible Properties

#### Load broadband program area shapefiles

```{r, cache=TRUE}
# {r, cache=TRUE}
map_program_area_rc <- st_read(file.path(geodatadir, "ReConnect R1 594 PFSA 05012020.shp")) %>%
  sf::st_transform(4326)
```

#### Create list of county FIPS codes that are within the program area for the database query

This dataset does not have a county FIPS code. Instead it provides a list of county names associated with that entry. We need to do some work to convert these lists of county names to a list of unique county fips codes.

```{r}
# {r}
int <- sf::st_intersects(map_program_area_rc, counties_us)

cnty_ints <- data.table(rc_rownum = numeric(), cntys_rownum = numeric(), geoid = character())
for (i in 1:length(int)) {
  u <- unlist(int[i])
  for (j in 1:length(int[[i]])) {
    geoid <- counties_us[u[j],]$GEOID
    newrow <- data.table(rc_rownum = i, cntys_rownum = u[j], geoid)
    cnty_ints <- rbindlist(list(cnty_ints, newrow))
    #print(paste(i, u[j]))
  }
}

cnty_ints_geoid <- cnty_ints[, .(geoid_cntys = paste(geoid, collapse = ", ")), rc_rownum]
map_program_area_rc$geoid_cnty <- ""

for (i in 1:nrow(cnty_ints_geoid)) {
  map_program_area_rc[cnty_ints_geoid[i,]$rc_rownum,]$geoid_cnty <- cnty_ints_geoid[i,]$geoid_cntys
}

geoid_cnty_unq <- unique(strsplit(paste(map_program_area_rc$geoid_cnty, collapse = ", "), ", ")[[1]])
geoid_cnty_unq <- geoid_cnty_unq[geoid_cnty_unq != ""]
counties_rc <- paste0("'", paste(geoid_cnty_unq, collapse = "','"), "'")

counties_rc_sql <- glue_sql(counties_rc)
```

#### Get property records for the program related counties from the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)
```

```{sql}
-- {sql connection=con, output.var="qry_res"}
SELECT geoid_cnty, p_id_iris_frmtd, property_centroid_longitude, property_centroid_latitude FROM corelogic_usda.current_tax_200627_latest_all_add_vars 
WHERE property_centroid_longitude IS NOT NULL 
AND geoid_cnty IN (?counties_rc_sql);
```

```{r}
# {r}
DBI::dbDisconnect(con)
```

#### Get intersection between program area shapefile and the selected properties

First convert database query result to an sf object.

```{r}
# {r}
db_rows_rc <- st_as_sf(qry_res, coords = c("property_centroid_longitude", "property_centroid_latitude"), crs = 4326)

int <- sf::st_intersects(db_rows_rc, map_program_area_rc)
int2 <- as.integer(as.character(int))
```

#### Add RUSID to intersecting properties

```{r}
# {r}
db_rows_rc$program <- "RC"
db_rows_rc$rusid <- as.character(map_program_area_rc$RUS_ID[unlist(int2)])
db_rows_rc <- db_rows_rc[!is.na(db_rows_rc$rusid),]
```

## BIP Program Eligible Properties

#### Load broadband program area shapefiles

```{r, cache=TRUE}
# {r, cache=TRUE}
map_program_area_bip <- st_read(file.path(geodatadir, "200409_BIP_ServAr_ID.shp")) %>%
  sf::st_transform(4326)
```

#### Create list of county FIPS codes that are within the program area for the database query

This dataset does not have a county FIPS code. Instead it provides a list of county names associated with that entry. We need to do some work to convert these lists of county names to a list of unique county fips codes.

```{r}
# {r}
int <- sf::st_intersects(map_program_area_bip, counties_us)

cnty_ints <- data.table(rc_rownum = numeric(), cntys_rownum = numeric(), geoid = character())
for (i in 1:length(int)) {
  u <- unlist(int[i])
  for (j in 1:length(int[[i]])) {
    geoid <- counties_us[u[j],]$GEOID
    newrow <- data.table(rc_rownum = i, cntys_rownum = u[j], geoid)
    cnty_ints <- rbindlist(list(cnty_ints, newrow))
  }
}

cnty_ints_geoid <- cnty_ints[, .(geoid_cntys = paste(geoid, collapse = ", ")), rc_rownum]
map_program_area_bip$geoid_cnty <- ""

for (i in 1:nrow(cnty_ints_geoid)) {
  map_program_area_bip[cnty_ints_geoid[i,]$rc_rownum,]$geoid_cnty <- cnty_ints_geoid[i,]$geoid_cntys
}

geoid_cnty_unq <- unique(strsplit(paste(map_program_area_bip$geoid_cnty, collapse = ", "), ", ")[[1]])
geoid_cnty_unq <- geoid_cnty_unq[geoid_cnty_unq != ""]
counties_bip <- paste0("'", paste(geoid_cnty_unq, collapse = "','"), "'")

counties_bip_sql <- glue_sql(counties_bip)
```

#### Get property records for the program related counties from the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)
```

```{sql}
-- {sql connection=con, output.var="qry_res"}
SELECT geoid_cnty, p_id_iris_frmtd, property_centroid_longitude, property_centroid_latitude FROM corelogic_usda.current_tax_200627_latest_all_add_vars 
WHERE property_centroid_longitude IS NOT NULL
AND property_centroid_latitude IS NOT NULL
AND geoid_cnty IN (?counties_bip_sql);
```

```{r}
# {r}
DBI::dbDisconnect(con)
```

#### Get intersection between program area shapefile and the selected properties

First convert database query result to an sf object.

```{r}
# {r}
db_rows_bip <- st_as_sf(qry_res, coords = c("property_centroid_longitude", "property_centroid_latitude"), crs = 4326)

int <- sf::st_intersects(db_rows_bip, map_program_area_bip)
int2 <- as.integer(as.character(int))
```

#### Add RUSID to intersecting properties

```{r}
# {r}
db_rows_bip$program <- "BIP"
db_rows_bip$rusid <- as.character(map_program_area_bip$RUS_ID[unlist(int2)])
db_rows_bip <- db_rows_bip[!is.na(db_rows_bip$rusid),]
```

## TCF Program Eligible Properties

#### Load broadband program area shapefiles

```{r, cache=TRUE}
# {r, cache=TRUE}
map_program_area_tcf <- st_read(file.path(geodatadir, "USDARD_RUS_TELCO_FARMBILL_06042020/USDARD_RUS_TELCO_FARMBILL_06042020.shp")) %>%
  sf::st_transform(4326)
```

#### Create list of county FIPS codes that are within the program area for the database query

This dataset does not have a county FIPS code. Instead it provides a list of county names associated with that entry. We need to do some work to convert these lists of county names to a list of unique county fips codes.

```{r}
# {r}
int <- sf::st_intersects(map_program_area_tcf, counties_us)

cnty_ints <- data.table(rc_rownum = numeric(), cntys_rownum = numeric(), geoid = character())
for (i in 1:length(int)) {
  u <- unlist(int[i])
  for (j in 1:length(int[[i]])) {
    geoid <- counties_us[u[j],]$GEOID
    newrow <- data.table(rc_rownum = i, cntys_rownum = u[j], geoid)
    cnty_ints <- rbindlist(list(cnty_ints, newrow))
    #print(paste(i, u[j]))
  }
}

cnty_ints_geoid <- cnty_ints[, .(geoid_cntys = paste(geoid, collapse = ", ")), rc_rownum]
map_program_area_tcf$geoid_cnty <- ""

for (i in 1:nrow(cnty_ints_geoid)) {
  map_program_area_tcf[cnty_ints_geoid[i,]$rc_rownum,]$geoid_cnty <- cnty_ints_geoid[i,]$geoid_cntys
}

geoid_cnty_unq <- unique(strsplit(paste(map_program_area_tcf$geoid_cnty, collapse = ", "), ", ")[[1]])
geoid_cnty_unq <- geoid_cnty_unq[geoid_cnty_unq != ""]
counties_tcf <- paste0("'", paste(geoid_cnty_unq, collapse = "','"), "'")

counties_tcf_sql <- glue_sql(counties_tcf)
```

#### Get property records for the program related counties from the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)
```

```{sql}
-- {sql connection=con, output.var="qry_res"}
SELECT geoid_cnty, p_id_iris_frmtd, property_centroid_longitude, property_centroid_latitude FROM corelogic_usda.current_tax_200627_latest_all_add_vars 
WHERE property_centroid_longitude IS NOT NULL 
AND property_centroid_latitude IS NOT NULL
AND geoid_cnty IN (?counties_tcf_sql);
```

```{r}
# {r}
DBI::dbDisconnect(con)
```

#### Get intersection between program area shapefile and the selected properties

First convert database query result to an sf object.

```{r}
# {r}
db_rows_tcf <- st_as_sf(qry_res, coords = c("property_centroid_longitude", "property_centroid_latitude"), crs = 4326)

int <- sf::st_intersects(db_rows_tcf, map_program_area_tcf)
int2 <- as.integer(as.character(int))
```

### Add RUSID to intersecting properties

```{r}
# {r}
db_rows_tcf$program <- "TCF"
db_rows_tcf$rusid <- as.character(map_program_area_tcf$RUS_ID[unlist(int2)])
db_rows_tcf <- db_rows_tcf[!is.na(db_rows_tcf$rusid),]
```

## TCI Program Eligible Properties

#### Load broadband program area shapefiles

```{r, cache=TRUE}
# {r, cache=TRUE}
map_program_area_tci <- st_read(file.path(geodatadir, "USDARD_RUS_TELCO_INFRA_06042020/USDARD_RUS_TELCO_INFRA_06042020.shp")) %>%
  sf::st_transform(4326)
```

#### Create list of county FIPS codes that are within the program area for the database query

This dataset does not have a county FIPS code. Instead it provides a list of county names associated with that entry. We need to do some work to convert these lists of county names to a list of unique county fips codes.

```{r}
# {r}
int <- sf::st_intersects(map_program_area_tci, counties_us)

cnty_ints <- data.table(rc_rownum = numeric(), cntys_rownum = numeric(), geoid = character())
for (i in 1:length(int)) {
  u <- unlist(int[i])
  for (j in 1:length(int[[i]])) {
    geoid <- counties_us[u[j],]$GEOID
    newrow <- data.table(rc_rownum = i, cntys_rownum = u[j], geoid)
    cnty_ints <- rbindlist(list(cnty_ints, newrow))
    #print(paste(i, u[j]))
  }
}

cnty_ints_geoid <- cnty_ints[, .(geoid_cntys = paste(geoid, collapse = ", ")), rc_rownum]
map_program_area_tci$geoid_cnty <- ""

for (i in 1:nrow(cnty_ints_geoid)) {
  map_program_area_tci[cnty_ints_geoid[i,]$rc_rownum,]$geoid_cnty <- cnty_ints_geoid[i,]$geoid_cntys
}

geoid_cnty_unq <- unique(strsplit(paste(map_program_area_tci$geoid_cnty, collapse = ", "), ", ")[[1]])
geoid_cnty_unq <- geoid_cnty_unq[geoid_cnty_unq != ""]
counties_tci <- paste0("'", paste(geoid_cnty_unq, collapse = "','"), "'")

counties_tci_sql <- glue_sql(counties_tci)
```

#### Get property records for the program related counties from the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)
```

```{sql}
-- {sql connection=con, output.var="qry_res"}
SELECT geoid_cnty, p_id_iris_frmtd, property_centroid_longitude, property_centroid_latitude FROM corelogic_usda.current_tax_200627_latest_all_add_vars 
WHERE property_centroid_longitude IS NOT NULL 
AND property_centroid_latitude IS NOT NULL
AND geoid_cnty IN (?counties_tci_sql);
```

```{r}
# {r}
DBI::dbDisconnect(con)
```

#### Get intersection between program area shapefile and the selected properties

First convert database query result to an sf object.

```{r}
# {r}
db_rows_tci <- st_as_sf(qry_res, coords = c("property_centroid_longitude", "property_centroid_latitude"), crs = 4326)

int <- sf::st_intersects(db_rows_tci, map_program_area_tci)
int2 <- as.integer(as.character(int))
```

#### Add RUSID to intersecting properties

```{r}
# {r}
db_rows_tci$program <- "TCI"
db_rows_tci$rusid <- as.character(map_program_area_tci$RUSID[unlist(int2)])
db_rows_tci <- db_rows_tci[!is.na(db_rows_tci$rusid),]
```

## Combine All Program Eligible Properties

```{r}
# {r}
combn <- rbindlist(list(db_rows_cc, db_rows_rc, db_rows_bip, db_rows_tcf, db_rows_tci))

combn_cast <- dcast(combn, geoid_cnty + p_id_iris_frmtd ~ program, value.var = "rusid")

saveRDS(combn_cast, file.path(geodatadir, "program_eligible_properties.RDS"))
```

## Write combined data to the database

```{r}
# {r}
con <- get_db_conn(db_host = "localhost", db_port = 5434)

dbWriteTable(con, c("corelogic_usda", "current_tax_200627_program_eligible_properties"), combn_cast, overwrite = TRUE, row.names = FALSE)

dbExecute(con, "ALTER TABLE corelogic_usda.current_tax_200627_program_eligible_properties ADD PRIMARY KEY (geoid_cnty, p_id_iris_frmtd)")

DBI::dbDisconnect(con)
```

## Join new table with current_tax_200627_latest_all_add_vars

```{sql}
-- {sql} * Run In psql/pgcli
SELECT a.*, b."BIP", b."CC", b."RC", b."TCF", b."TCI"
INTO corelogic_usda.current_tax_200627_latest_all_add_vars_add_progs
FROM corelogic_usda.current_tax_200627_latest_all_add_vars a
LEFT JOIN corelogic_usda.current_tax_200627_program_eligible_properties b
ON a.geoid_cnty = b.geoid_cnty AND a.p_id_iris_frmtd = b.p_id_iris_frmtd;

ALTER TABLE corelogic_usda.current_tax_200627_latest_all_add_vars_add_progs ADD PRIMARY KEY (geoid_cnty, p_id_iris_frmtd);
```
