---
title: "Import Data File"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

------------------------------------------------------------------------

## DOWNLOAD 2014 DATA

```{r}
# url <- "http://transition.fcc.gov/form477/BroadbandData/Fixed/Dec14/Version%203/US-Fixed-without-Satellite-Dec2014.zip"
# temp <- tempfile()
# download.file(url,temp)
# unlink(temp)

# grab temp pathname and go intoto terminal 
# ran this to see name of file in zip 
# unzip -l /tmp/RtmpwK6I3G/file33274b60567b
# ran this to unzip from terminal:
# unzip /tmp/RtmpwK6I3G/file33274b60567b -d /project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/

fcc_2014 <- read.csv("/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fbd_us_without_satellite_dec2014_v3.csv")
```

## NEW VARIABLES:

-   Get the maximum speed per census block where data is available in
    FCC data
-   Compare that maximum speed to our eligibility threshold

Note that we tried the following thresholds for 2014.

-   10/1
-   4/1
-   4/768
-   3 (upload + download) \*\* selected

```{r include = FALSE}
# fcc_2014_block_max <- fcc_2014 %>%
#   group_by(StateAbbr, BlockCode) %>%
#   summarise(MaxAdDown_ = max(MaxAdDown),
#             MaxAdUp_ = max(MaxAdUp)) %>%
#   mutate(speed_elig_down = ifelse(MaxAdDown_ >= 10, 0, 1),
#          speed_elig_up = ifelse(MaxAdUp_ >= 1, 0, 1),
#          speed_inelig = ifelse(MaxAdDown_ >= 10 & MaxAdUp_ >= 1, 1, 0),
#          speed_elig = ifelse(MaxAdDown_ >= 10 & MaxAdUp_ >= 1, 0, 1))

# #threshold of 4/1
# fcc_2014_block_max <- fcc_2014 %>%
#   group_by(StateAbbr, BlockCode) %>%
#   summarise(MaxAdDown_ = max(MaxAdDown),
#             MaxAdUp_ = max(MaxAdUp)) %>%
#   mutate(speed_elig_down = ifelse(MaxAdDown_ >= 4, 0, 1),
#          speed_elig_up = ifelse(MaxAdUp_ >= 1, 0, 1),
#          speed_inelig = ifelse(MaxAdDown_ >= 4 & MaxAdUp_ >= 1, 1, 0),
#          speed_elig = ifelse(MaxAdDown_ >= 4 & MaxAdUp_ >= 1, 0, 1))

# #threshold of 4/768
# fcc_2014_block_max_ <- fcc_2014 %>%
#   group_by(StateAbbr, BlockCode) %>%
#   summarise(MaxAdDown_ = max(MaxAdDown),
#             MaxAdUp_ = max(MaxAdUp)) %>%
#   mutate(speed_elig_down = ifelse(MaxAdDown_ >= 4, 0, 1),
#          speed_elig_up = ifelse(MaxAdUp_ >= 0.768, 0, 1),
#          speed_inelig = ifelse(MaxAdDown_ >= 4 & MaxAdUp_ >= 0.768, 1, 0),
#          speed_elig = ifelse(MaxAdDown_ >= 4 & MaxAdUp_ >= 0.768, 0, 1))

# saveRDS(fcc_2014_block_max, "/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fcc_2014_block_maxspdelig.RDS")
# saveRDS(fcc_2014_block_max, "/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fcc_2014_block_maxspdelig_threshold_4_1.RDS")
# saveRDS(fcc_2014_block_max_, "/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fcc_2014_block_maxspdelig_threshold_4_768.RDS")

```

```{r}
# #threshold of 3 mbps (download + upload)
fcc_2014_block_max_ <- fcc_2014 %>%
  group_by(StateAbbr, BlockCode) %>%
  summarise(MaxAdDown_ = max(MaxAdDown),
            MaxAdUp_ = max(MaxAdUp)) %>%
  mutate(total_speed = MaxAdDown_ + MaxAdUp_,
         speed_elig_down = NA, #ifelse(MaxAdDown_ >= 4, 0, 1),
         speed_elig_up = NA, #ifelse(MaxAdUp_ >= 0.768, 0, 1),
         speed_inelig = ifelse(total_speed < 3, 1, 0),
         speed_elig = ifelse(total_speed >= 3, 0, 1))

saveRDS(fcc_2014_block_max_, "/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fcc_2014_block_maxspdelig_threshold_total3.RDS")

```

```{r}

```


## DOWNLOAD 2018 DATA

# url \<- "<https://www.fcc.gov/form477/BroadbandData/Fixed/Dec18/Version%202/US-Fixed-without-Satellite-Dec2018.zip>"

# temp \<- tempfile()

# download.file(url,temp)

# unlink(temp)

# grab temp pathname and go intoto terminal

# ran this to see name of file in zip

# unzip -l /tmp/RtmpSdiLsM/file360be78486be5

# ran this to unzip from terminal:

# unzip /tmp/RtmpSdiLsM/file360be78486be5 -d /project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/

## MANIPULATE 2018 DATA

fcc_2018 \<-
read.csv("/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fbd_us_without_satellite_dec2018_v2.csv")

fcc_2018_block_max \<- fcc_2018 %\>% group_by(StateAbbr, BlockCode) %\>%
summarise(MaxAdDown\_ = max(MaxAdDown), MaxAdUp\_ = max(MaxAdUp)) %\>%
mutate(speed_elig_down = ifelse(MaxAdDown\_ \>= 10, 0, 1), speed_elig_up
= ifelse(MaxAdUp\_ \>= 1, 0, 1), speed_inelig = ifelse(MaxAdDown\_ \>=
10 & MaxAdUp\_ \>= 1, 1, 0), speed_elig = ifelse(MaxAdDown\_ \>= 10 &
MaxAdUp\_ \>= 1, 0, 1))

# saveRDS(fcc_2018_block_max, "/project/biocomplexity/sdad/projects_data/project_data/usda/rural_broadband/fcc_availability/fcc_2018_block_maxspdelig.RDS")
